---
title: "Visualization Exercises"
author: "Ethan Marcano"
date: '2022-03-21'
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
library(formatR)
library(tidyverse)
library(igraph)
library(maps)
library(ggmap)
library(wesanderson)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE, echo = TRUE)
```
# Visualization Exercises

## Question 1: Visualization Attributes of Parole Violators

### Part A)

```{r parole_read, message=FALSE}
Parole <- read_csv("Parole.csv")
```

```{r parole_plot}
ggplot(data = Parole, aes(x = Age)) + 
  geom_histogram(binwidth = 5, color = "blue")
```

i)
+ The age bracket with the most parolees is the 25-30 age bracket. The one with the least parolees is the 65-69 age bracket.
ii)
+ It creates a blue outline for each bracket, enabling the user to see the data more clearly.

### Part B)

```{r parole_gender}
ggplot(data = Parole, aes(x = Age)) + geom_histogram(binwidth = 5) + facet_grid(Male ~ .)
```

i)
+ The majority of prisoners in all age brackets are male. However the male prisoners are on the younger side, while the female prisoners trend towards middle age.

```{r parole_gender_flipped}
ggplot(data = Parole, aes(x = Age)) + 
  geom_histogram(binwidth = 5) + facet_grid(. ~ Male)
```

ii)
+ It flips the axis for the graph.

```{r parole_fill}
ggplot(data = Parole, aes(x = Age, fill = Male)) + 
  geom_histogram(binwidth = 5, position = "identity", alpha = 0.5)
```

iii)
+ The faceting histograms look better in delineating between different data groups, and enables easy comparison.

### Part C)

```{r parole_month}
ggplot(data = Parole, aes(x = TimeServed)) + 
  geom_histogram(binwidth = (1), color = "red")
```

i)
+ The most common length of time served is 5 years.

```{r parole_month_decimal}
ggplot(data = Parole, aes(x = TimeServed)) + 
  geom_histogram(binwidth = 0.1, color = "red")
```

ii)
+ The most common length of time served is 3 years. It may be necessary to make visualizations more precise as to avoid interpretation errors.

```{r parole_facet_crime}
ggplot(data = Parole, aes(x = TimeServed)) +
  geom_histogram(binwidth = 1, color = "red") +
  facet_grid(. ~ Crime)
```

iii)
+ Out of the named crimes, a drug arrest suggests a longer time served. Other crimes (may referring to more severe misdemeanors or felonies) also tend to carry a harder penalty.

```{r parole_fill_crime}
ggplot(data = Parole, aes(x = TimeServed, fill = Crime)) +
  geom_histogram(binwidth = 1, position = "identity", alpha = 0.5)
```
+ The faceting histogram enables you to look at all types of data, and I prefer for much of the same reasons as the gender histograms.

## Question 2: Visualizing Network Data

### Part A)

```{r users_ read, message=FALSE}
users <- read_csv("Users.csv")
edges <- read_csv("Edges.csv")
```

```{r users_count}
nrow(users)
```
i) 
+ There are 59 users in our dataset.

```{r users_average}
nrow(edges)/nrow(users)
```
ii)
+ There is an average of about 2.5 friends between users in the network.

```{r users_locale}
locale_count <- users %>%
  count(Locale)
locale_count
```
iii)
+ Locale B is the most common, outnumbering A by 44 users.

```{r users_all_gender_school}
gender_school <- users %>%
  count(School, Gender)
gender_school
```

iv)
+ Neither school is an all-girls or all-boys school.

### Part B)

```{r users_igraph}
g <- graph.data.frame(edges, FALSE, users)
plot(g, vertex.size = 5, vertex.label = NA)
```

i)
+ There are three connected components outside the main group.
ii)
+ There are seven users with no friends.

```{r users_degrees}
degree(g)
```

iii)
+ There are nine users with 10 or more friends.

### Part C)
```{r users_gender_colors}
V(g)$color[V(g)$Gender == "A"] = "red"
V(g)$color[V(g)$Gender == "B"] = "blue"

plot(g, vertex.size = 5, vertex.label = NA)
```

i)
+ It seems that users of Gender B tend to have a high number of friends and make up a vast majority of connections.

```{r users_locale_colors}
V(g)$color[V(g)$School == "A"] = "red"
V(g)$color[V(g)$School == "B"] = "blue"
V(g)$color[V(g)$School == "AB"] = "yellow"

plot(g, vertex.size = 5, vertex.label = NA)
```

ii)
+ Yes, those users are both friends with each other.

## Question 3: Election Forecasting

### Part A)

```{r polling_read}
polling <- read_csv("Polling.csv")
```

```{r polling_map_colors}
pal <- wes_palette("Zissou1", 100, type = "continuous")
```

```{r polling_map_plot_setup}
statesMap <- map_data("state")
electionMap <- merge(statesMap, polling, by.x = "region", by.y = "State")
electionMap <- electionMap[order(electionMap$order),]
```

```{r polling_map_plot}
ggplot(electionMap, aes(x = long, y = lat, group = group, fill = DiffCount)) +
  geom_polygon(color="black")
```

```{r polling_map_scale}
ggplot(electionMap, aes(x = long, y = lat, group = group, fill = DiffCount)) +
  geom_polygon(color="black") + scale_fill_gradientn(colours = pal)
```

+ Missouri, Montana, and Florida stick out as consistent polling states. Ohio is a notable outlier.

### Part B)

```{r polling_map_surveyUSA}
ggplot(electionMap, aes(x = long, y = lat, group = group, fill = SurveyUSA)) +
  geom_polygon(color="black") + 
  scale_fill_gradient(low = "blue", high = "red")
```

```{r polling_map_rasmussen}
ggplot(electionMap, aes(x = long, y = lat, group = group, fill = Rasmussen)) +
  geom_polygon(color="black") + 
  scale_fill_gradient(low = "blue", high = "red")
```

i)
+ The most notable thing that while both groups share some common states between them, SurveyUSA focuses on primarily on the Mid-Atlantic region while Rasmussen covers the Midwest. Rasmussen also trends centrist, while SurveyUSA trends Republican. Compared to the DiffCount map, both groups predict a redder Midwest than DiffCount.

ii)
+ All three maps predict a tight race, leaning Republican. For the 2012 election, none of these maps are correct.

### Part C)
```{r polling_map_republican}
ggplot(electionMap, aes(x = long, y = lat, group = group, fill = Republican)) +
  geom_polygon(color="black") + 
  scale_fill_gradient(low = "blue", high = "red")
```

iii)
+ All of the maps got a purple Midwest relatively right, as well as a blue West Coast and Northeast.
+ If I were to create a predictive model, I would stick to aggregating different polling sources, particularly in swing state regions such as the Midwest and Mid-Atlantic.